Hypertension and microbiome

### 2019-05-14 articletwo.Rmd
Better salt analysis

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(finriskmetagcommon)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(ggpmisc)
library(Maaslin)
library(parallel)
library(officer)
library(flextable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
```

@@

```{r core taxa}
pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
```

@@

```{r lacto data frame}
lacto.abund <- subset_taxa(pseq.genus, Genus == "Lactobacillus") %>%
    abundances(., transform = "clr") %>%
    t %>%
    as.data.frame %>%
    tibble::rownames_to_column("id") %>%
    setNames(tolower(replace.brackets(names(.))))

lacto.full <- meta(pseq.genus) %>%
    tibble::rownames_to_column("id") %>%
    full_join(., lacto.abund, by = "id") %>%
    dplyr::mutate(fake_salt = relevel(factor(dplyr::ntile(BL_AGE, 4)), ref='1'),
                  lacto_groups = relevel(factor(dplyr::ntile(lactobacillus.bacteria, 4)), ref='1'))
```

```{r lacto adjusted means}
lacto.lm.salt <- lm(lactobacillus.bacteria ~ fake_salt + BL_AGE + SEX, data=lacto.full)

lacto.mean.unadjusted <- data.frame(
    fake_salt = as.factor(seq(4)),
    SEX = as.factor(0),
    BL_AGE = mean(lacto.full %>% pull(BL_AGE)))

lacto.mean.adjusted <- predict(lacto.lm.salt, lacto.mean.unadjusted, interval = "confidence") %>%
    as.data.frame %>%
    dplyr::rename(estimate = fit,
           conf.low = lwr,
           conf.high = upr) %>%
    tibble::rownames_to_column("term")
```

```{r lacto odds for hypertension}
lacto.odds <- glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX,
                  family = binomial(link=logit),
                  data=lacto.full) %>%
    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
    filter(!grepl("Intercept", term)) %>%
    select(term, estimate, conf.low, conf.high) %>%
    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%
    dplyr::filter(grepl("lacto", term))
```

```{r lacto lm residuals}
ggplot2::autoplot(lm(lactobacillus.bacteria ~ fake_salt + BL_AGE + SEX, data=lacto.full))
```

```{r lacto logistic residuals}
ggplot2::autoplot(glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX, family = binomial(link=logit), data=lacto.full))
```

```{r lacto anova}
lacto.anova <- aov(lactobacillus.bacteria ~ fake_salt, data = lacto.full)
summary(lacto.anova)
TukeyHSD(lacto.anova)$fake_salt %>% kable
```

```{r lactoto test normality}
grid.arrange(ggqqplot(lacto.full$lactobacillus.bacteria, title = "Lactobacillus"),
             ggqqplot(lacto.full$BL_AGE, title = "BL_AGE"), ncol=2)
```

```{r forest plot lacto}
myforestplot <- function(data, xlab = "", ylab = "", ylimits = c(0,10), interceptone = FALSE, subtitle = "") {
    ggplot(data = data,
           aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
        geom_point() +
        { if (interceptone) geom_hline(aes(fill=fake_salt), yintercept=1, linetype=2) } +
        geom_errorbar(width=0.1)+ 
        ggplot2::theme_minimal()  +
        theme(plot.title=element_text(size=16),
              axis.title=element_text(size=12),
              strip.background = element_blank()) +
        theme(panel.border = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              axis.ticks.x = element_line(),
              axis.ticks.y = element_line(),
              axis.line = element_line(colour = "black")) +
        scale_x_discrete(labels=c("Q1", "Q2", "Q3", "Q4")) +
        scale_y_continuous(limits = ylimits, breaks = pretty(ylimits, 4)) + 
        xlab(xlab) +
        ylab(ylab) +
        labs(subtitle = subtitle)
}

gsalt <- grid.arrange(
    myforestplot(lacto.mean.adjusted,
                 xlab = "Dietary salt",
                 ylab = "Lactobacillus abundance",
                 subtitle = "A.",
                 ylimits = c(2.5,3)),
    myforestplot(lacto.odds,
                 xlab = "Lactobacillus abundance",
                 ylab = "Odds of hypertension",
                 subtitle = "B.",
                 ylimits = c(0.5, 1.5),
                 interceptone = TRUE),
    ncol=2)
ggsave(file = "cache/salt-forest.png", plot=gsalt, height=4, width=6, dpi = 300)
```

>>>>

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(Maaslin)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
```

@@

@@

Load genus data with dU-NA and combine it in a data frame with clr transformed lactobacillus abundances.

```{r lacto data frame}
pseq.genus.salt <- import_salt_pseq("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")

abund.genus.salt <- subset_taxa(pseq.genus.salt, Genus == "Lactobacillus") %>%
    abundances(., transform = "clr") %>%
    t %>%
    as.data.frame %>%
    tibble::rownames_to_column("id") %>%
    setNames(tolower(replace.brackets(names(.))))

df.genus.salt <- meta(pseq.genus.salt) %>%
    tibble::rownames_to_column("id") %>%
    full_join(., abund.genus.salt, by = "id") %>%
    dplyr::mutate(salt_groups = relevel(factor(dplyr::ntile(dUNA, 3)), ref='1'),
                  lacto_groups = relevel(factor(dplyr::ntile(lactobacillus.bacteria, 3)), ref='1'))
```

### Daily urine natrium

Associations between lactobacillus and dU-NA

```{r abund salt lm continuous}
lm.lacto.duna <- lm(lactobacillus.bacteria ~ dUNA + BL_AGE + SEX, data = df.genus.salt)
lm.lacto.duna %>% tidy %>% kable
```

### Salt groups

Association between lactobacillus and salt groups

```{r abund salt lm group}
lm.lacto.saltgroup <- lm(lactobacillus.bacteria ~ salt_groups + BL_AGE + SEX, data = df.genus.salt)
lm.lacto.saltgroup %>% tidy %>% kable
```

Assesment for linear model asuumptions

```{r abund salt lm group checks plot, fig.width=12, fig.height=8}
par(mfrow=c(3,3))
plot(gvlma::gvlma(lm.lacto.saltgroup), onepage = FALSE)
```

Anova between lactobacillus and salt groups

```{r salt anova}
aov.lacto.saltgroup <- aov(lactobacillus.bacteria ~ salt_groups + BL_AGE + SEX, data = df.genus.salt)
aov.lacto.saltgroup %>% xtable %>% kable
```

Tukey's range test for lactobacillus and salt groups

```{r salt HSD}
hsd.lacto.saltgroup <- TukeyHSD(aov.lacto.saltgroup, "salt_groups")
hsd.lacto.saltgroup$salt_groups %>% kable
```

### Adjusted means

Calculate adjusted means for lactobacillus

```{r lacto adjusted means}
df.genus.salt.unadjusted <- with(df.genus.salt, data.frame(salt_groups = as.factor(seq(3)),
                                                           SEX = as.factor(0),
                                                           BL_AGE = mean(BL_AGE)))

adjmean.lacto.saltgroup <- predict(lm.lacto.saltgroup,
                               df.genus.salt.unadjusted,
                               interval = "confidence") %>%
    as.data.frame %>%
    dplyr::rename(estimate = fit,
                  conf.low = lwr,
                  conf.high = upr) %>%
    tibble::rownames_to_column("term")
```

### Odds for hypertension

Odds for hypertension

```{r glm htn lacto}
glm.hypertension.lactogroup <- glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX,
                  family = binomial(link=logit),
                  data = df.genus.salt) %>%
    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
glm.hypertension.lactogroup %>% kable
```

```{r odds htn lacto}
odds.hypertension.lactogroup <- glm.hypertension.lactogroup %>%
    filter(!grepl("Intercept", term)) %>%
    select(term, estimate, conf.low, conf.high) %>%
    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%
    dplyr::filter(grepl("lacto", term))
```

### Twin-forest plot

```{r forest plot lacto}
gsalt <- grid.arrange(
    myforestplot(adjmean.lacto.saltgroup,
                 xlab = "Dietary salt",
                 ylab = "Lactobacillus abundance",
                 subtitle = "A.",
                 my_y_scale = scale_y_continuous(limits = c(2.5,3), breaks = seq(2.5, 3.0, 0.1))),
    myforestplot(odds.hypertension.lactogroup,
                 xlab = "Lactobacillus abundance",
                 ylab = "Odds of hypertension",
                 subtitle = "B.",
                 my_y_scale = scale_y_log10(limits = c(0.5, 2), breaks = seq(0.5, 2.0, 0.5)),
                 interceptone = TRUE),
    ncol=2)

ggsave(file = "cache/salt-forest.png", plot=gsalt, height=4, width=6, dpi = 300)
```
